---
title: "Spotify analysis"
output:
  html_document: default
  html_notebook: default
---

```{r}
library(spotifyr)
library(tidyverse)
```

Set your keys to system environment variables
```{r}
Sys.setenv(SPOTIFY_CLIENT_ID = 'e22cc372a8c74c419517806f9e14c715')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '7bd8a8f3705741eb8581ddd4fe0cc70a')
```
What artist are you looking at?
```{r}
artist <- "Arctic Monkeys"
```



Get artist uri
```{r}
uris <- get_artists(artist_name = artist) %>% 
  filter(artist_name == artist)  # filter because there's an "Arctic Monkeys tribute band" on Spotify that shows up in results

uri <- uris$artist_uri[1]
```

Get songs

```{r}
songs <- get_album_tracks(get_albums(artist_uri = uri))
```

Get songs and attributes for the songs by joining the song dataframe to attribute dataframes on the track_uri

```{r}
songs <- songs %>% 
  inner_join(get_track_popularity(songs)) %>% 
  inner_join(get_track_audio_features(songs)) %>%
  arrange(-track_popularity)

# join to albums to get dates for album
songs <- songs %>% 
  inner_join(get_albums(uri))
```
#### Theme
```{r}
spotify_theme <- theme(legend.position = "none",
                  panel.background = element_rect(fill = "white"),
                  panel.grid.major = element_line(color = "white"),
                  panel.grid.minor = element_line(color = "white"),
                  panel.border = element_rect(color = "lightgrey", fill = NA))
```


#### Analysis

Scatterplot matrix

```{r}
# only numeric variables for scatterplots
num_var <- songs %>% 
  select(track_popularity, danceability, energy, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms)

pairs(num_var)
```

Linear regression.

```{r}
model <- lm(track_popularity ~ danceability + tempo + speechiness + acousticness + valence + duration_ms, data = songs)

summary(model)
```

#### Valence (measure of cheerfulness)

```{r}
pairs(num_var[, c("valence", "duration_ms", "track_popularity", "tempo")])
```
Danceability
```{r}
songs %>% 
  select(track_name, track_popularity, danceability) %>% 
  arrange(-danceability)
```

Valence vs popularity
```{r}
songs %>% 
  ggplot(aes(valence, track_popularity)) +
  geom_jitter(size = .5) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_text(aes(label = track_name), size = 3, angle = 10) +
  spotify_theme +
  labs(x = "Song's Valence Score",
       y = "Song's Popularity on Spotify",
       title = "Valence vs. Popularity")
```

Danceability vs popularity
```{r}
songs %>% 
  ggplot(aes(danceability, track_popularity)) +
  geom_jitter(size = .5) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_text(aes(label = track_name), size = 3, angle = 10) +
  spotify_theme +
  labs(x = "Song's Danceability Score",
       y = "Song's Popularity on Spotify",
       title = "Danceability vs. Popularity")
```

Plotting multiple histograms at once using https://www.r-bloggers.com/quick-plot-of-all-variables/

Gather all variables and facet wrap by value

```{r}
songs %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~key, scales = "free") +
  geom_histogram()
```

Trying to decide which variables to cluster by. Perhaps Danceability, Energy, Tempo?
```{r}
pairs(num_var[, c("danceability", "energy", "tempo", "valence", "duration_ms", "acousticness")])
```

IDEA: Gif of movement across albums

```{r}
songs %>% 
  ggplot(aes(tempo, valence, colour = album_name)) +
  geom_point()
```
Mean tempo / valence per album
```{r}
songs %>% 
  group_by(album_name) %>% 
  summarise(average_tempo = mean(tempo), average_valence = mean(valence), average_energ = mean(energy), average_dance = mean(danceability)) %>% 
  ggplot(aes(average_dance, average_valence, colour = album_name)) +
  geom_point() +
  spotify_theme +
  theme(legend.position = "none") +
  geom_text(aes(label = album_name),
            position = position_nudge(.00, .002),
            ) +
  labs(x = "Average Danceability",
       y = "Average Cheerfulness",
       title = "Arctic Monkeys Albums' Danceability and Cheerfulness",
       subtitle = "Using Spotify Data")
```
